Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-22 checking-executions.tex
modify

We explain the proof of this property on the execution $e$ in \figurename~\ref{fig:concurrent execution for EPQ1}(a) where $p_1 \prec p$, $p_1 \prec p_2$, and the predicate $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds. Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$. Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{rm}(z_2)$, $\textit{put}(x_3,p_2)$, and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_2)-w$), and $v$ be any sequence containing the remaining operations. In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return actions, called \emph{linearization point}. The order between the linearization points defines the sequence $l'$. \figurename~\ref{fig:concurrent execution for EPQ1}(a) draws linearization points for the operations in $e \setminus x$ defining the linearization $l'$~\footnote{In general, there may exist multiple ways of choosing linearization points to define the same linearization. Our construction is agnostic to this choice.}.
%(the order of these linearization points defines the sequence $l'$). 
We show how to construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3\in\seqPQ$ such that $e \sqsubseteq l''$. 
%In this example, . We also explicitly draw the linearization points according to $l'$. We construct $l''$ as follows, and this method does not rely on the locations of linearization points of $l'$ and operations of $x$.
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] An operation is called $p$-comparable (resp., $p$-incomparable) when it receives as argument a value of priority comparable to $p$ (resp., incomparable to $p$). We could try to define $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ to the operations in $u$, $v$ and $w$, respectively. However, this is incorrect, since $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ imposes no restriction on  $p$-incomparable operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ to $p$-incomparable operations in $u \cdot v$ is correct. In this example, this projection is $\textit{put}(z_1,p_3) \cdot \textit{rm}(z_2)$ which is incorrect.

@@

\item[-] The second step is to define $U'$ and $V'$. $U'$ contains two kinds of operations: (1) operations whose linearization points are before $\textit{ret}(\textit{put},x)$, and (2) other $\textit{put}$ operations with priority $p$. $V'$ contains the remaining operations. In this example, $U'$ contains $\textit{put}(z_1,p_3)$ and $\textit{put}(x_2,p_2)$.

>>>>

We explain the proof of this property on the execution $e$ in \figurename~\ref{fig:concurrent execution for EPQ1}(a) where $p_1 \prec p$, $p_1 \prec p_2$, and the predicate $\mathsf{Has\text{-}MatchedMaxPriority}(e)$ holds. Assume that there exist two sequential executions $l$ and $l'$ such that $e \sqsubseteq l=u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w$, $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ holds, and $e \setminus x \sqsubseteq l' \in \seqPQ$. Let $u=\epsilon$, $w$ be any sequence containing the set of operations $\textit{put}(z_2,p_2)$, $\textit{put}(x_3,p)$ and $\textit{rm}(z_1)$ (we distinguish them by adding the suffix $-w$ to their name, e.g., $\textit{rm}(z_1)-w$), and $v$ be any sequence containing the remaining operations. In general, the linearization $l'$ can be defined by choosing for each operation, a point in time between its call and return actions, called \emph{linearization point}. The order between the linearization points defines the sequence $l'$. \figurename~\ref{fig:concurrent execution for EPQ1}(a) draws linearization points for the operations in $e \setminus x$ defining the linearization $l'$~\footnote{In general, there may exist multiple ways of choosing linearization points to define the same linearization. Our construction is agnostic to this choice.}.
%(the order of these linearization points defines the sequence $l'$).
We show how to construct a sequence $l''= l''_1 \cdot \textit{put}(x,p) \cdot l''_2 \cdot \textit{rm}(x) \cdot l''_3\in\seqPQ$ such that $e \sqsubseteq l''$.
%In this example, . We also explicitly draw the linearization points according to $l'$. We construct $l''$ as follows, and this method does not rely on the locations of linearization points of $l'$ and operations of $x$.
\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] An operation is called $p$-comparable (resp., $p$-incomparable) when it receives as argument a value of priority comparable to $p$ (resp., incomparable to $p$). We could try to define $l''_1$, $l''_2$ and $l''_3$ as the projection of $l'$ to the operations in $u$, $v$ and $w$, respectively. However, this is incorrect, since $\mathsf{MatchedMaxPriority\text{-}Seq}(l,x)$ imposes no restriction on  $p$-incomparable operations in $u \cdot v$, and thus, there is no guarantee that the projection of $l'$ to $p$-incomparable operations in $u \cdot v$ is correct. In this example, this projection is $\textit{put}(z_1,p_2) \cdot \textit{rm}(z_2)$ which is incorrect.

@@

\item[-] The second step is to define $U'$ and $V'$. $U'$ contains two kinds of operations: (1) operations whose linearization points are before $\textit{ret}(\textit{put},x,p)$, and (2) other $\textit{put}$ operations with priority $p$. $V'$ contains the remaining operations. In this example, $U'$ contains $\textit{put}(z_1,p_2)$ and $\textit{put}(x_2,p)$.
